---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# osgeo0

<!-- badges: start -->
<!-- badges: end -->

R interface to the GDAL Geographic Network Model (GNM) via Python's `osgeo.gnm` and [reticulate](https://rstudio.github.io/reticulate/).

> **Status**: alpha — API is exploratory and will change.

## What is this?

osgeo0 wraps `osgeo.gnm` as the first module in a broader R interface to the Python `osgeo.*` bindings. GNM was chosen as the starting point because it has a small, well-defined API surface that exercises the full lifecycle patterns (create, populate, analyse, close) needed for all future `osgeo.*` wrapping.

The package uses:

- **[wk](https://paleolimbot.github.io/wk/)** for geometry interchange (WKB bytes across the Python–R bridge)
- **[S7](https://rconsortium.github.io/S7/)** for the class system
- **[reticulate](https://rstudio.github.io/reticulate/)** to call Python's `osgeo` modules

## Installation

```r
# install.packages("pak")
pak::pak("mdsumner/osgeo0")
```

You also need a Python environment with GDAL's Python bindings:

```bash
conda install -c conda-forge gdal python
```

## Quick start

```{r quick-start}
reticulate::use_python("/usr/bin/python3")

library(osgeo0)

# Create a network
net <- gnm_create(tempdir(), "water_network", srs = "EPSG:4326")
net


```

## Design

See [`inst/design/osgeo0-design-v2.md`](inst/design/osgeo0-design-v2.md) for the full design document.

Key patterns that will carry forward to `osgeo.ogr`, `osgeo.gdal`, etc.:

- S7 value class + private environment for mutable Python dataset references
- OGR ↔ wk bridge via WKB bytes (no competing `libgdal` linkage)
- `reg.finalizer()` for automatic cleanup on GC
- Eager materialisation of results (no dangling Python references in returned objects)

## License

MIT


## Code of Conduct
  
Please note that the osgeo0 project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
